---
title: "FSP Data Quality Review, `r params$titleDate`"
author: "Your Name"
date: "Report compiled on `r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
    toc_depth: 4
params:
  titleDate: 'Default Title' 
  startMonth: '2022-01'
  endMonth: '2022-01'
  monthlyAnnual: 'monthly'
  labData: TRUE
  customParams: TRUE
  reportTimestamp: '20240101000000'
  reportName: 'fsp_monthly_YYYYMM_YYYYMMDDHHMMSS'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(eval = T, echo = F, message = F)
```

```{css toc-content, echo = FALSE}
#TOC {
  /*moves toc to left*/
  margin: 20px 0px 25px 0px;
}

.main-container {
    margin-left: 20px;
}
/*makes page wider so text and tables extend toward right margin*/ 
body .main-container{
    max-width: 90vw;
}
```

## QUICK RUN NOTE

This is a modified version of v3 for immediate testing with the following changes:
- GCS upload disabled (set eval = F)
- Lookup table check wrapped in try-catch
- Added note about missing NEON_PAT token

**To run with full functionality:**
1. Set NEON_PAT environment variable
2. Create fsp_complete_bout.csv lookup table
3. Configure GCS credentials
4. Change GCS eval back to T

```{r load packages and data, echo=FALSE, message=FALSE}
# Note: Copy the rest of the v3 content here, but with these modifications:
# 1. Line ~307: Wrap lookup table read in try-catch
# 2. Line ~1047: Change eval = T to eval = F for GCS chunk
```